---
title: "Anscombe's triptychs"
author: "Ruben Arslan"
date: "22/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```


We attach the tidyverse packages to simulate the data and the brms package.

```{r}
library(tidyverse)
library(brms)
```

## Basic model
We first generate the predictors, M and X for 900 rows.
We set a seed, so that the simulated random numbers are reproducible.
The following assumptions are made:

- `M` takes one of three values (0, 1, 2) that are equally probable.
- `X` depends on M to some degree (a coefficient of `0.8`) and varies substantially in addition to that.
- `X` has a standard deviation of ~1.
- We already simulate the residual variation for the outcome variable `Y`.


```{r}
cases <- list()
n <- 900
set.seed(20191005)
moderation <-  tibble(
  M = sample(0:2, n, replace = TRUE),
  M_f = factor(M, 0:2, c("low", "medium", "high")),
  X = 0.8 * (M-1) + rnorm(n, sd = 0.77),
  residual = rnorm(n)
)
sd(moderation$M)
sd(moderation$X)
```


## "Ideal" case
Here, we simulate the "ideal" moderation scenario that most people will assume, i.e. there is a causal interaction between X and M, on top of main effects for both X and M. The interaction takes the form of an attenuation of the main effect for X (or, respectively, M).
```{r}
moderation <- moderation %>% 
  mutate(
  Y = -0.66 + 0.1*X + 0.5*M + 0.3 * X * M + 0.69 * residual
)
sd(moderation$X)
sd(moderation$Y)
```

```{r}
cases$assumed <- moderation
round(coef(lm(Y ~ X*M, cases$assumed)),2)

ggplot(cases$assumed, 
       aes(X, Y)) +
  geom_point(aes(color = M), alpha = 0.1) +
  geom_smooth(method = "lm") +
  scale_color_viridis_c(guide = F) +
  facet_wrap(~ M_f)
```

## Floor effect
```{r}
moderation <-  cases$assumed %>% 
  mutate(
      Y = -1.1 + 0.68*X + 0.8*M + 0.95 * residual,
      Y = -0.66 + if_else(Y < -0.7, -0.7, Y) + 0.52
  )
sd(moderation$Y)
round(coef(lm(Y ~ X*M, moderation)),2)
round(coef(lm(Y ~ X*M, cases$assumed)),2)
```

```{r}
cases$floor <- moderation

ggplot(cases$floor, 
       aes(X, Y)) +
  geom_point(aes(color = M), alpha = 0.1) +
  geom_smooth(method = "lm") +
  scale_color_viridis_c(guide = F) +
  facet_wrap(~ M_f)
```


## Residual variance
```{r}
moderation <-  cases$assumed %>% 
  mutate(
      Y = -0.66 + 0.1*X + 0.5*M + 0.3 * X * M + 0.49*((exp(M)/3-0.2))*residual
  )
sd(moderation$Y)
mean(moderation$Y)
round(coef(lm(Y ~ X*M, moderation)),2)
round(coef(lm(Y ~ X*M, cases$assumed)),2)
```

```{r}
cases$residual <- moderation

moderation %>% group_by(M_f) %>% 
  do(broom::tidy(cor.test(.$X, .$Y)))
moderation %>% group_by(M_f) %>% 
  do(broom::tidy(cor.test(.$X, resid(lm(.$Y ~ .$M)))))

# plot(lm(Y ~ X*M, moderation))
round(coef(lm(Y ~ X*M, moderation)),2)

ggplot(cases$residual, 
       aes(X, Y)) +
  geom_point(aes(color = M), alpha = 0.1) +
  geom_smooth(method = "lm") +
  scale_color_viridis_c(guide = F) +
  facet_wrap(~ M_f)
```

## Triptychs
```{r}
options(digits = 2)
options(pillar.sigfig = 3)


model <- list()
model$assumed <- brm(Y ~ X*M_f, data = cases$assumed, cores = 4, backend = "cmdstanr", file = "assumed")
model$floor <- brm(Y ~ X*M_f, data = cases$floor, cores = 4, backend = "cmdstanr", file = "floor")
model$residual <- brm(Y ~ X*M_f, data = cases$residual, cores = 4, backend = "cmdstanr", file = "residual")

fixed_model <- list()
fixed_model$assumed <- brm(Y ~ X*M_f, data = cases$assumed, cores = 4, backend = "cmdstanr", file = "assumed")
fixed_model$floor <- brm(Y | cens(censored) ~ X*M_f, data = cases$floor %>% mutate(censored = if_else(Y < -0.83, "left", "none")), cores = 4, backend = "cmdstanr", file = "fixed_floor")
fixed_model$residual <- brm(bf(Y ~ X*M_f, sigma ~ M_f), data = cases$residual, cores = 4, backend = "cmdstanr", file = "fixed_residual")

library(bayesplot)

library(cowplot)

cases$assumed[,c("estimate__", "std.error__", "lower__", "upper__")] <- fitted(model$assumed)

cases$floor[,c("estimate__", "std.error__", "lower__", "upper__")] <- fitted(model$floor)

cases$residual[,c("estimate__", "std.error__", "lower__", "upper__")] <- fitted(model$residual)


three_cases <- bind_rows(cases, .id = "case")
three_cases %>% select(-M_f, -ends_with("__")) %>% group_by(case) %>% summarise_all(list(~mean(.), ~sd(.)))
three_cases %>% group_by(case) %>% do(broom::tidy(lm(Y ~ X*M, .))) %>% select(case, term, estimate) %>% tidyr::pivot_wider(id_cols = case, names_from = term, values_from = estimate)

ggplot(three_cases, 
       aes(X, Y)) +
  # geom_point(aes(color = M), alpha = 0.1) +
  geom_line(aes(X, estimate__)) +
  scale_color_viridis_c(guide = F) +
  facet_grid(case ~ M_f) +
  theme_minimal()

```


```{r}
(cors <- three_cases %>% group_by(case, M_f) %>%
  do(broom::tidy(cor.test(.$X, .$Y))))
ggplot(three_cases, 
       aes(X, Y)) +
  geom_point(aes(color = M), alpha = 0.1) +
  geom_smooth(aes(X, estimate__, ymin = lower__, ymax = upper__), 
              color = "black", stat = 'identity') +
  scale_color_viridis_c(guide = F, begin = 0.2, end = 0.8) +
  facet_grid(case ~ M_f, labeller = labeller(
    case = c("assumed" = '"ideal" case', 
             "floor" = "floor effect,\nno latent interaction",
             "residual" = "smaller correlation\nat largest slope"),
    M_f = c("low" = "Moderator: low",
            "medium" = "Moderator: medium",
            "high" = "Moderator: high"))) +
  geom_text(aes(label = sprintf('italic("r")==%.2f', estimate, conf.low, conf.high)), x= -2.5, y=2.5, data = cors, hjust = 0, parse = TRUE, size = 3) +
  ggthemes::theme_clean() +
  coord_cartesian(ylim = c(-2.5,3.6))

ggsave("anscombe_triptychs.pdf", width = 6, height = 6)
ggsave("anscombe_triptychs.png", width = 6, height = 6)
```


## Posterior predictive checks
```{r}

plot_grid(
  pp_check(model$assumed) + guides(color = F), # pp_check(model$assumed, type = "error_scatter_avg", alpha = 0.1),
  pp_check(model$floor) + theme(legend.position = c(1,0.9), legend.justification = "right"), # pp_check(model$floor, type = "error_scatter_avg", alpha = 0.1), 
  pp_check(model$residual) + guides(color = F),
#  pp_check(model$residual, type = "error_scatter_avg", alpha = 0.1),
  nrow = 1, labels = "AUTO", label_size = 6, axis = "lb", align = "hv")
ggsave("three_dgms_ppc.pdf", width = 5, height = 2)
ggsave("three_dgms_ppc.png", width = 5, height = 2)

plot_grid(
  pp_check(fixed_model$assumed) + guides(color = F), #pp_check(fixed_model$assumed, type = "error_scatter_avg", alpha = 0.1),
  pp_check(fixed_model$floor) + theme(legend.position = c(1,0.9), legend.justification = "right"),# pp_check(fixed_model$floor, type = "error_scatter_avg", alpha = 0.1), 
  pp_check(fixed_model$residual) + guides(color = F),
 # pp_check(fixed_model$residual, type = "error_scatter_avg", alpha = 0.1),
  nrow = 1, labels = "AUTO", label_size = 6, axis = "lb", align = "hv")
ggsave("fixed3_dgms_ppc.pdf", width = 5, height = 2)
ggsave("fixed3_dgms_ppc.png", width = 5, height = 2)

```


