# Differences in slopes or residual variance

```{r}
library(tidyverse)
n <- 50
days_per_person <- 300
n_days <- n*days_per_person
library(brms)
set.seed(20191005)
```

## Model 1: Main effect only

### Data-generating model
```{r}
people <- tibble(
  id = 1:n,
  single = if_else(runif(n) > 0.5, 1, 0),
  single_f = factor(single, 0:1, c("in relationship", "single")),
  average_job_satisfaction = rnorm(n),
  average_relationship_satisfaction = rnorm(n),
  average_well_being = rnorm(n, if_else(single == 1, -0.5, 0.5))
)
diary <-  people %>% full_join(
  tibble(
    id = rep(1:n, each = days_per_person),
  ), by = 'id') %>% 
  mutate(
    Job_satisfaction = rnorm(n_days, average_job_satisfaction),
    Relationship_satisfaction = rnorm(n_days, average_relationship_satisfaction),
    OWB = Job_satisfaction + 3 * rnorm(n_days, average_well_being)
  )
round(cor(diary %>% select(single, Job_satisfaction, Relationship_satisfaction, OWB)),2)
```

### Inspect correlations
```{r}
cor.test(diary$Job_satisfaction, diary$OWB)
diary %>% group_by(single_f) %>%
  summarise(cor(Job_satisfaction, OWB))
```

### Scatter plots
```{r}
ggplot(diary, aes(Job_satisfaction, OWB)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm")
ggplot(diary, aes(Job_satisfaction, OWB)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  facet_wrap(~ id)

cors <- diary %>% group_by(single_f) %>%
  do(broom::tidy(cor.test(.$Job_satisfaction, .$OWB)))

ggplot(diary, aes(Job_satisfaction, OWB, color = single_f)) +
  facet_wrap(~ single_f) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  scale_color_discrete(guide = F)+
  geom_text(aes(label = sprintf("%.2f [%.2f;%.2f]", estimate, conf.low, conf.high)), x= 0, y=-13, data = cors) +
  theme_bw()
```


### Models
```{r}
m_simple <- brm(bf(OWB ~ Job_satisfaction + (1|id)), data = diary, 
               cores = 4, iter = 2000, file = "models/dgm_main/m_simple", 
               backend = "cmdstanr") %>%
  add_criterion("loo")

m_sigma <- brm(bf(OWB ~ Job_satisfaction + (1|id),
                sigma ~ single), data = diary, 
               cores = 4, iter = 2000, file = "models/dgm_main/m_sigma", 
               backend = "cmdstanr") %>%
  add_criterion("loo")

m_int <- brm(bf(OWB ~ Job_satisfaction * single) + (1|id), data = diary, 
             cores = 4, iter = 2000, file = "models/dgm_main/m_int",
                 backend = "cmdstanr") %>%
  add_criterion("loo")

m_sigma_int <- brm(bf(OWB ~ Job_satisfaction * single + (1|id),
                     sigma ~ single), data = diary, 
                 cores = 4, iter = 2000, file = "models/dgm_main/m_sigma_int",
                 backend = "cmdstanr") %>%
  add_criterion("loo")
```

### Inferences
```{r}
knitr::kable(fixef(m_int))
knitr::kable(fixef(m_sigma))
loo_compare(m_simple, m_int)
loo_compare(m_simple, m_sigma)
loo_compare(m_simple, m_int, m_sigma, m_sigma_int)
```

## Model 2: Difference in residual variance, but not slope

### Data-generating model
```{r}
diary <-  diary %>% 
  mutate(
    OWB = Job_satisfaction + (1-single) * Relationship_satisfaction + 3 * rnorm(n_days, average_well_being)
  )
round(cor(diary %>% select(single, Job_satisfaction, Relationship_satisfaction, OWB)),2)
```

### Inspect correlations
```{r}
cor.test(diary$Job_satisfaction, diary$OWB)
diary %>% group_by(single_f) %>%
  summarise(cor(Job_satisfaction, OWB))
```

### Scatter plots
```{r}
ggplot(diary, aes(Job_satisfaction, OWB)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm")

cors <- diary %>% group_by(single_f) %>%
  do(broom::tidy(cor.test(.$Job_satisfaction, .$OWB)))

ggplot(diary, aes(Job_satisfaction, OWB, color = single_f)) +
  facet_wrap(~ single_f) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  scale_color_discrete(guide = F)+
  geom_text(aes(label = sprintf("%.2f [%.2f;%.2f]", estimate, conf.low, conf.high)), x= 0, y=-13, data = cors) +
  theme_bw()
```



### Models
```{r}
m_simple <- brm(bf(OWB ~ Job_satisfaction + (1|id)), data = diary, 
               cores = 4, iter = 2000, file = "models/dgm_sig/m_simple", 
               backend = "cmdstanr") %>%
  add_criterion("loo")

m_sigma <- brm(bf(OWB ~ Job_satisfaction + (1|id),
                sigma ~ single), data = diary, 
               cores = 4, iter = 2000, file = "models/dgm_sig/m_sigma", 
               backend = "cmdstanr") %>%
  add_criterion("loo")

 m_int <- brm(bf(OWB ~ Job_satisfaction * single) + (1|id), data = diary, 
             cores = 4, iter = 2000, file = "models/dgm_sig/m_int",
                 backend = "cmdstanr") %>%
  add_criterion("loo")

m_sigma_int <- brm(bf(OWB ~ Job_satisfaction * single + (1|id),
                     sigma ~ single), data = diary, 
                 cores = 4, iter = 2000, file = "models/dgm_sig/m_sigma_int",
                 backend = "cmdstanr") %>%
  add_criterion("loo")
```

### Inferences
```{r}
knitr::kable(fixef(m_int))
knitr::kable(fixef(m_sigma))
loo_compare(m_simple, m_int)
loo_compare(m_simple, m_sigma)
loo_compare(m_simple, m_int, m_sigma, m_sigma_int)
```


## Model 3: Difference in residual variance and slope cancel out in correlation

### Data-generating model
```{r}
n <- 2000
diary <- tibble(
  close_friend = if_else(runif(n) > 0.5, 1, 0),
  months_known = rpois(n, 5 + 3*close_friend),
  trait = rnorm(n),
  close_friend_f = factor(close_friend, 0:1, c("acquaintance", "close friend")),
  perception = (1 + close_friend) * trait + (1 + 2*close_friend) * rnorm(n)
)
round(cor(diary %>% select(-close_friend_f)),2)
```

### Inspect correlations
```{r}
cor.test(diary$trait, diary$perception)
diary %>% group_by(close_friend) %>%
  summarise(cor(trait, perception))
diary %>% group_by(months_known) %>%
  summarise(cor(trait, perception), n())
```

### Scatter plots
```{r}
ggplot(diary, aes(trait, perception)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm")

(cors <- diary %>% group_by(close_friend_f) %>%
  do(broom::tidy(cor.test(.$trait, .$perception))))
slopes <- diary %>% group_by(close_friend_f) %>%
  do(broom::tidy(lm(.$perception ~ .$trait), conf.int = TRUE)[2,])

ggplot(diary, aes(trait, perception, color = close_friend_f)) +
  facet_wrap(~ close_friend_f, labeller = labeller(close_friend_f = c("acquaintance"= "Moderator 0","close friend" = "Moderator 1"))) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  scale_color_discrete(guide = F)+
  geom_text(aes(label = sprintf("r=%.2f [%.2f;%.2f]", estimate, conf.low, conf.high)), x= -2, y=-9, data = cors) +
  geom_text(aes(label = sprintf("b=%.2f [%.2f;%.2f]", estimate, conf.low, conf.high)), x= -2, y=-10, data = slopes) +
  xlab("Var X") +
  ylab("Var Y") +
  theme_bw()

ggplot(diary, aes(trait, perception)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~ months_known)
  # geom_smooth(method = "lm")
```


### Models
```{r}
m_simple <- brm(perception ~ trait, data = diary, 
               cores = 4, iter = 2000, file = "models/dgm_sig_int_nocor/m_simple", 
               backend = "cmdstanr") %>%
  add_criterion("loo")

m_sigma <- brm(bf(perception ~ trait,
                sigma ~ close_friend + months_known), data = diary, 
               cores = 4, iter = 2000, file = "models/dgm_sig_int_nocor/m_sigma", 
               backend = "cmdstanr") %>%
  add_criterion("loo")

m_int <- brm(bf(perception ~ trait * (close_friend + months_known)), data = diary, 
             cores = 4, iter = 2000, file = "models/dgm_sig_int_nocor/m_int",
                 backend = "cmdstanr") %>%
  add_criterion("loo")

m_sigma_int <- brm(bf(perception ~ trait * (close_friend),
                     sigma ~ close_friend), data = diary, 
                 cores = 4, iter = 2000, file = "models/dgm_sig_int_nocor/m_sigma_int",
                 backend = "cmdstanr") %>%
  add_criterion("loo")
```

### Inferences
```{r}
knitr::kable(fixef(m_int))
knitr::kable(fixef(m_sigma))
knitr::kable(fixef(m_sigma_int))
loo_compare(m_simple, m_int)
loo_compare(m_simple, m_sigma)
loo_compare(m_simple, m_int, m_sigma, m_sigma_int)
```



```{r}
diary <- tibble(
  close_friend = if_else(runif(n) > 0.8, 1, 0),
  months_known = rpois(n, 5 + 3*close_friend),
  trait = rnorm(n),
  perception = (close_friend + 1)* trait + (1 + close_friend) * rnorm(1000)
)
round(cor(diary),2)

cor.test(diary$trait, diary$perception)
diary %>% group_by(close_friend) %>%
  summarise(cor(trait, perception))
diary %>% group_by(months_known) %>%
  summarise(cor(trait, perception), n())

ggplot(diary, aes(trait, perception)) +
  geom_point() +
  geom_smooth(method = "lm")

diary$close_friend_f <- factor(diary$close_friend, 0:1, c("acquaintance", "close friend"))

cors <- diary %>% group_by(close_friend_f) %>%
  do(broom::tidy(cor.test(.$trait, .$perception)))

ggplot(diary, aes(trait, perception, color = close_friend_f)) +
  facet_wrap(~ close_friend_f) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  scale_color_discrete(guide = F)+
  geom_text(aes(label = sprintf("%.2f [%.2f;%.2f]", estimate, conf.low, conf.high)), x= 0, y=-13, data = cors) +
  theme_bw()
```


## Cohen's
```{r}
library(effsize)
n <- 5000

prepost <- tibble(
  id = 1:n,
  treatment = rep(1:2, each = n/2),
  t1 = rnorm(n),
  t2 = t1 - treatment + rnorm(n, 0, 1.5)
)
sd(prepost$t2)
cohen.d(t2 ~ treatment, prepost)
cohen.d(t2-t1 ~ treatment, prepost)

prepost <- tibble(
  id = 1:n,
  treatment = rep(1:2, each = n/2),
  t1 = rnorm(n),
  t2 = t1 - treatment + rnorm(n, 0, treatment)
)
sd(prepost$t2)
cohen.d(t2 ~ treatment, prepost)
cohen.d(t2-t1 ~ treatment, prepost)
ggplot(prepost, aes(t1, t2, color = treatment)) +
  geom_point(alpha = 0.1) +
  geom_pointrange(aes(0, t2), fun.data = "mean_se", stat = "summary") +
  geom_smooth(method = 'lm') +
  geom_abline(intercept = 0, slope = 1, linetype = 'dashed') +
  facet_wrap(~ treatment)
```

